---
title: "Reproducible Research: Peer Assessment 1"
author: "Wouter van Amsterdam"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data


```{r load, results='hide', cache = T}
library(data.table)
activity <- fread("activity.csv")
```


## What is mean total number of steps taken per day?


```{r steps_day, message=F}
library(data.table)
library(dplyr)
act_day <- activity[, list(steps = sum(steps, na.rm = T)), by = "date"]
act_day[, mean(steps)]
invisible(act_day[, hist(steps)]) # suppress histogram details
act_day[, list(mean_steps = mean(steps), median_steps = median(steps))]
```


## What is the average daily activity pattern?

```{r steps_pattern}
act_interval <- activity[, list(steps = mean(steps, na.rm = T)), by = "interval"]
act_interval[, plot(steps~interval, type = 'l')]
act_interval[steps == max(steps), list(interval, steps)]
```



## Imputing missing values


Look at missing values. Create copy of data.table. 
Impute values based on average value for that interval
```{r missings}
colSums(is.na(activity))
act_imp <- copy(activity)
act_imp[is.na(steps), steps:=act_interval[interval == interval, steps]]
colSums(is.na(act_imp))
```


Based on inputed data, create histogram again and median and mean values. The histogram changed clearly. 
```{r imputed_summary}
act_imp_day <- act_imp[, list(steps = sum(steps)), by = 'date']
invisible(act_imp_day[, hist(steps)])
act_imp_day[, list(mean_steps = mean(steps), median_steps = median(steps))]
```

Mean increased by 
`r round(act_imp_day[, mean(steps)] - act_day[, mean(steps)], 0)` 
steps, quite dramatic. The median only increased by 
`r round(act_imp_day[, median(steps)] - act_day[, median(steps)], 0)` 
steps, which is better.


## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays}
act_imp[, weekend:=factor(weekdays(lubridate::ymd(date)) %in% c("Saturday", "Sunday"), levels = c(T,F), labels = c("weekend", "weekday"))]
act_imp[, table(weekend)]

act_imp_weekend <- act_imp[, list(steps = mean(steps)), by = 'weekend,interval']

library(lattice)
xyplot(steps ~ interval | weekend, data = act_imp_weekend, type = 'l')
```

From the look of it, this guy walks to work, or to a coffee shop each workday.






